### 2024-03-25 docs/projects/geography/chicago-bikes/index.qmd
Chicago bikes - rerendered after bugfix

::: {.sourceClojure}
```clojure
(ns index
  (:require [tablecloth.api :as tc]
            [clojure.math :as math]
            [tech.v3.datatype.datetime :as datetime]
            [clojure.string :as str]
            [scicloj.noj.v1.vis.hanami :as hanami]
            [aerial.hanami.templates :as ht]
            [fastmath-clustering.core :as clustering]
            [geo
             [geohash :as geohash]
             [jts :as jts]
             [spatial :as spatial]
             [io :as geoio]
             [crs :as crs]]
            [scicloj.kindly.v4.kind :as kind]
            [scicloj.kindly.v4.api :as kindly])
  (:import (org.locationtech.jts.geom Geometry Point Polygon Coordinate)
           (org.locationtech.jts.geom.prep PreparedGeometry
                                           PreparedLineString
                                           PreparedPolygon
                                           PreparedGeometryFactory)))
```
:::

@@

For metric computations we need to convert them to a coordinate system
which is locally correct in terms of distances in a region around Chicago:
NAD83 / Illinois East
[EPSG:26971](https://epsg.io/26971-1742)

@@

::: {.sourceClojure}
```clojure
(defn lat-lng->local-coords [lat lng]
  (-> (geo.jts/coordinate -89.27 37.06)
      geo.jts/point
      wgs84->Chicago
      geo.jts/coord
      ((fn [^Coordinate coord]
         [(.getX coord)
          (.getY coord)]))))
```
:::

@@

TODO

>>>>

::: {.sourceClojure}
```clojure
(ns index
  (:require [tablecloth.api :as tc]
            [clojure.math :as math]
            [tech.v3.datatype.datetime :as datetime]
            [tech.v3.datatype.functional :as fun]
            [clojure.string :as str]
            [scicloj.noj.v1.vis.hanami :as hanami]
            [aerial.hanami.templates :as ht]
            [fastmath-clustering.core :as clustering]
            [geo
             [geohash :as geohash]
             [jts :as jts]
             [spatial :as spatial]
             [io :as geoio]
             [crs :as crs]]
            [scicloj.kindly.v4.kind :as kind]
            [scicloj.kindly.v4.api :as kindly])
  (:import (org.locationtech.jts.geom Geometry Point Polygon Coordinate)
           (org.locationtech.jts.geom.prep PreparedGeometry
                                           PreparedLineString
                                           PreparedPolygon
                                           PreparedGeometryFactory)))
```
:::

@@

For metric computations (e.g., k-means clustering),
we need to convert them to a coordinate system
which is locally correct in terms of distances in a region around Chicago:
NAD83 / Illinois East
[EPSG:26971](https://epsg.io/26971-1742)

@@

::: {.sourceClojure}
```clojure
(defn lat-lng->local-coords [lat lng]
  (-> (geo.jts/coordinate lat lng)
      geo.jts/point
      wgs84->Chicago
      geo.jts/coord
      ((fn [^Coordinate coord]
         [(.getX coord)
          (.getY coord)]))))
```
:::

@@

::: {.sourceClojure}
```clojure
(-> processed-trips
    (tc/add-column :local-distance #(fun/sqrt
                                     (fun/+ (fun/sq (fun/- (:start-local-x %)
                                                           (:end-local-x %)))
                                            (fun/sq (fun/- (:start-local-y %)
                                                           (:end-local-y %)))))))
```
:::

::: {.clay-dataset}
data/kaggle-cyclistic/202304_divvy_tripdata.csv.gz [426155 21]:

|         :ride_id | :rideable_type |         :started_at |           :ended_at |        :start_station_name | :start_station_id |         :end_station_name | :end_station_id |  :start_lat |   :start_lng |    :end_lat |     :end_lng | :member_casual | :hour |                        :start-local-coords |                          :end-local-coords |  :start-local-x |  :start-local-y |    :end-local-x |    :end-local-y | :local-distance |
|------------------|----------------|---------------------|---------------------|----------------------------|-------------------|---------------------------|-----------------|------------:|-------------:|------------:|-------------:|----------------|------:|--------------------------------------------|--------------------------------------------|----------------:|----------------:|----------------:|----------------:|----------------:|
| 8FE8F7D9C10E88C7 |  electric_bike | 2023-04-02T08:37:28 | 2023-04-02T08:41:37 |                            |                   |                           |                 | 41.80000000 | -87.60000000 | 41.79000000 | -87.60000000 |         member |     8 |  [506107.9325773436 -1.4235352994997954E7] | [506137.42192310654 -1.4235316240199374E7] | 506107.93257734 | -1.42353530E+07 | 506137.42192311 | -1.42353162E+07 |     47.12257137 |
| 34E4ED3ADF1D821B |  electric_bike | 2023-04-19T11:29:02 | 2023-04-19T11:52:12 |                            |                   |                           |                 | 41.87000000 | -87.65000000 | 41.93000000 | -87.68000000 |         member |    11 |  [501519.8754807859 -1.4231908162239876E7] |  [498721.7934556093 -1.4229901392452776E7] | 501519.87548079 | -1.42319082E+07 | 498721.79345561 | -1.42299014E+07 |   3443.31061597 |
| 5296BF07A2F77CB5 |  electric_bike | 2023-04-19T08:41:22 | 2023-04-19T08:43:22 |                            |                   |                           |                 | 41.93000000 | -87.66000000 | 41.93000000 | -87.66000000 |         member |     8 | [500470.98093146156 -1.4231382653142933E7] | [500470.98093146156 -1.4231382653142933E7] | 500470.98093146 | -1.42313827E+07 | 500470.98093146 | -1.42313827E+07 |      0.00000000 |
| 40759916B76D5D52 |  electric_bike | 2023-04-19T13:31:30 | 2023-04-19T13:35:09 |                            |                   |                           |                 | 41.92000000 | -87.65000000 | 41.91000000 | -87.65000000 |         member |    13 |  [501374.92347210384 -1.423208764589309E7] |  [501403.9254779691 -1.4232051758343631E7] | 501374.92347210 | -1.42320876E+07 | 501403.92547797 | -1.42320518E+07 |     46.14144070 |
| 77A96F460101AC63 |  electric_bike | 2023-04-19T12:05:36 | 2023-04-19T12:10:26 |                            |                   |                           |                 | 41.91000000 | -87.65000000 | 41.91000000 | -87.63000000 |         member |    12 |  [501403.9254779691 -1.4232051758343631E7] |  [503154.9287078667 -1.4233533354956321E7] | 501403.92547797 | -1.42320518E+07 | 503154.92870787 | -1.42335334E+07 |   2293.71768835 |
| 8D6A2328E19DC168 |  electric_bike | 2023-04-19T12:17:34 | 2023-04-19T12:21:38 |                            |                   |                           |                 | 41.91000000 | -87.63000000 | 41.92000000 | -87.65000000 |         member |    12 |  [503154.9287078667 -1.4233533354956321E7] |  [501374.92347210384 -1.423208764589309E7] | 503154.92870787 | -1.42335334E+07 | 501374.92347210 | -1.42320876E+07 |   2293.14049610 |
| C97BBA66E07889F9 |  electric_bike | 2023-04-19T09:35:48 |    2023-04-19T09:45 |                            |                   |                           |                 | 41.93000000 | -87.66000000 | 41.91000000 | -87.65000000 |         member |     9 | [500470.98093146156 -1.4231382653142933E7] |  [501403.9254779691 -1.4232051758343631E7] | 500470.98093146 | -1.42313827E+07 | 501403.92547797 | -1.42320518E+07 |   1148.07983018 |
| 6687AD4C575FF734 |  electric_bike | 2023-04-11T16:13:43 | 2023-04-11T16:18:41 |                            |                   |                           |                 | 42.00000000 | -87.66000000 | 41.99000000 | -87.66000000 |         member |    16 |  [500268.5489835212 -1.4231632572144713E7] | [500297.48516240803 -1.4231596883177325E7] | 500268.54898352 | -1.42316326E+07 | 500297.48516241 | -1.42315969E+07 |     45.94567272 |
| A8FA4F73B22BC11F |  electric_bike | 2023-04-11T16:29:24 | 2023-04-11T16:40:23 |                            |                   |                           |                 | 41.99000000 | -87.66000000 | 42.00000000 | -87.66000000 |         member |    16 | [500297.48516240803 -1.4231596883177325E7] |  [500268.5489835212 -1.4231632572144713E7] | 500297.48516241 | -1.42315969E+07 | 500268.54898352 | -1.42316326E+07 |     45.94567272 |
| 81E158FE63D99994 |  electric_bike | 2023-04-19T17:35:40 | 2023-04-19T17:36:11 |                            |                   |                           |                 | 41.88000000 | -87.65000000 | 41.88000000 | -87.65000000 |         member |    17 |   [501490.896685202 -1.4231944068150206E7] |   [501490.896685202 -1.4231944068150206E7] | 501490.89668520 | -1.42319441E+07 | 501490.89668520 | -1.42319441E+07 |      0.00000000 |
|              ... |            ... |                 ... |                 ... |                        ... |               ... |                       ... |             ... |         ... |          ... |         ... |          ... |            ... |   ... |                                        ... |                                        ... |             ... |             ... |             ... |             ... |             ... |
| BAE25DBA0CD0DF40 |  electric_bike | 2023-04-14T07:12:34 | 2023-04-14T07:42:07 | Richmond St & Diversey Ave |             15645 | Clark St & Ida B Wells Dr |    TA1305000009 | 41.93186915 | -87.70120764 | 41.87593267 | -87.63058454 |         member |     7 |  [496862.65843494714 -1.422833795268237E7] | [503203.23756304523 -1.4233366677138694E7] | 496862.65843495 | -1.42283380E+07 | 503203.23756305 | -1.42333667E+07 |   8092.65181120 |
| 5A98F86A573AAB2C |  electric_bike | 2023-04-24T07:27:02 | 2023-04-24T07:58:22 | Richmond St & Diversey Ave |             15645 | Clark St & Ida B Wells Dr |    TA1305000009 | 41.93198085 | -87.70128024 | 41.87593267 | -87.63058454 |         member |     7 |  [496855.9967031725 -1.4228332970829802E7] | [503203.23756304523 -1.4233366677138694E7] | 496855.99670317 | -1.42283330E+07 | 503203.23756305 | -1.42333667E+07 |   8100.96696311 |
| 92B02F2FBDC1FB9F |   classic_bike | 2023-04-12T08:16:48 | 2023-04-12T08:40:08 |     Michigan Ave & Lake St |      TA1305000011 |        May St & Taylor St |           13160 | 41.88602200 | -87.62439800 | 41.86948210 | -87.65548640 |         member |     8 |  [503715.7528127164 -1.4233861403126204E7] | [501040.95186117536 -1.4231500346004825E7] | 503715.75281272 | -1.42338614E+07 | 501040.95186118 | -1.42315003E+07 |   3567.79355636 |
| 119A6C53607EAA4A |  electric_bike | 2023-04-28T07:24:54 | 2023-04-28T07:53:44 | Richmond St & Diversey Ave |             15645 | Clark St & Ida B Wells Dr |    TA1305000009 | 41.93190396 | -87.70125461 | 41.87593267 | -87.63058454 |         member |     7 |  [496858.4550664372 -1.4228334598338569E7] | [503203.23756304523 -1.4233366677138694E7] | 496858.45506644 | -1.42283346E+07 | 503203.23756305 | -1.42333667E+07 |   8098.02951217 |
| 9BCF1E8BA027EAFA |  electric_bike | 2023-04-21T07:15:06 | 2023-04-21T07:41:45 | Richmond St & Diversey Ave |             15645 | Clark St & Ida B Wells Dr |    TA1305000009 | 41.93193305 | -87.70126295 | 41.87593267 | -87.63058454 |         member |     7 |  [496857.6431128598 -1.4228334082633112E7] | [503203.23756304523 -1.4233366677138694E7] | 496857.64311286 | -1.42283341E+07 | 503203.23756305 | -1.42333667E+07 |   8098.98613308 |
| A17D800CE963661A |   classic_bike | 2023-04-11T15:46:42 | 2023-04-11T15:50:03 |     Michigan Ave & Lake St |      TA1305000011 |    Clark St & Randolph St |    TA1305000030 | 41.88602200 | -87.62439800 | 41.88457623 | -87.63188991 |         member |    15 |  [503715.7528127164 -1.4233861403126204E7] |  [503063.6502967095 -1.4233301315734675E7] | 503715.75281272 | -1.42338614E+07 | 503063.65029671 | -1.42333013E+07 |    859.61362107 |
| 8B441A6C436E9900 |   classic_bike | 2023-04-29T21:20:21 | 2023-04-29T21:30:19 |       Halsted St & 18th St |             13099 | Blue Island Ave & 18th St |           13135 | 41.85750568 | -87.64599144 | 41.85755600 | -87.66153500 |         casual |    21 |  [501907.1861189178 -1.4232159853149524E7] | [500545.77541704464 -1.4231010242258897E7] | 501907.18611892 | -1.42321599E+07 | 500545.77541704 | -1.42310102E+07 |   1781.86539868 |
| 3980D64BE11540F1 |   classic_bike | 2023-04-24T09:16:05 | 2023-04-24T09:22:27 |       Halsted St & 18th St |             13099 | Blue Island Ave & 18th St |           13135 | 41.85750568 | -87.64599144 | 41.85755600 | -87.66153500 |         casual |     9 |  [501907.1861189178 -1.4232159853149524E7] | [500545.77541704464 -1.4231010242258897E7] | 501907.18611892 | -1.42321599E+07 | 500545.77541704 | -1.42310102E+07 |   1781.86539868 |
| 3EF4B49FF7DAA02C |   classic_bike | 2023-04-18T07:53:51 | 2023-04-18T07:59:16 | Franklin St & Jackson Blvd |      TA1305000025 |    Clark St & Randolph St |    TA1305000030 | 41.87770796 | -87.63532114 | 41.88457623 | -87.63188991 |         casual |     7 |  [502783.1163597462 -1.4233022413426103E7] |  [503063.6502967095 -1.4233301315734675E7] | 502783.11635975 | -1.42330224E+07 | 503063.65029671 | -1.42333013E+07 |    395.58284532 |
| 210B2ED6583DC231 |   classic_bike | 2023-04-29T07:33:55 | 2023-04-29T07:38:57 |     Michigan Ave & Lake St |      TA1305000011 |    Clark St & Randolph St |    TA1305000030 | 41.88602200 | -87.62439800 | 41.88457623 | -87.63188991 |         casual |     7 |  [503715.7528127164 -1.4233861403126204E7] |  [503063.6502967095 -1.4233301315734675E7] | 503715.75281272 | -1.42338614E+07 | 503063.65029671 | -1.42333013E+07 |    859.61362107 |
| D29CB39B9E3FC46A |  electric_bike | 2023-04-18T08:00:32 | 2023-04-18T08:02:35 | Franklin St & Jackson Blvd |      TA1305000025 | Clark St & Ida B Wells Dr |    TA1305000009 | 41.87813417 | -87.63525450 | 41.87593267 | -87.63058454 |         casual |     8 | [502787.71168951405 -1.4233028887672516E7] | [503203.23756304523 -1.4233366677138694E7] | 502787.71168951 | -1.42330289E+07 | 503203.23756305 | -1.42333667E+07 |    535.50301123 |
